---
title: "Customer Transaction"
author: "Ayodeji Yekeen"
date: "12/31/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# ANZ Customer Transaction Data Analysis

This analysis is based on a synthesised transaction dataset containing 3 months worth of transactions for 100 hypothetical customers. It contains purchases, recurring transactions, and salary transactions.

The dataset is designed to simulate realistic transaction behaviours that are observed in ANZ’s real transaction data.

### load required libraries
```{r}
library(stringr)
library(lubridate)
library(tidyverse)
library(modelr)
library(sp)
library(leaflet)
library(geosphere)
library(knitr)
library(rpart)
```
## Exploratory data analysis

### Read the transaction dataset
```{r}
df <- read.csv('ANZ_synthesised_transaction_dataset.csv')
```

```{r}
#list of column names
colnames(df)

# how many rows are in the dataframe?
nrow(df)

# what is the size (column, row) of the dataframe?
dim(df)

# see list columns and data types
str(df)
```

```{r}
# view first 6 rows of the dataframe
head(df)
```

## Data Summary 
```{r}
# statistical summary of data
summary(df)
```
### Date Cleaning & Manipulation
```{r}
# change the format of date column
df$date<- as.Date(df$date,format = "%d/%m/%Y")

# derive weekday and hour data of each transaction
df$extraction = as.character(df$extraction)
df$hour = hour(as.POSIXct(substr(df$extraction,12,19),format="%H:%M:%S"))
df$weekday = weekdays(df$date)

# the dateset only contain records for 91 days, one day is missing
DateRange <- seq(min(df$date), max(df$date), by = 1)
DateRange[!DateRange %in% df$date]
```
### confirm the one -to -one link of account_id and customer_id
```{r}
df %>% select(account,customer_id) %>%
  unique() %>%
  nrow()
```
### split customer & merchant lat_long into individual columns for analysis
```{r}
dfloc = df[,c("long_lat","merchant_long_lat")]
dfloc<- dfloc %>% separate("long_lat", c("c_long", "c_lat"),sep=' ')
dfloc<- dfloc %>% separate("merchant_long_lat", c("m_long", "m_lat"),sep=' ')
dfloc<- data.frame(sapply(dfloc, as.numeric))
df <- cbind(df,dfloc)
```
### check the range of customer location
```{r}

# filtering out transactions for those who don't reside in Australia
df_temp <- df %>%
  filter (!(c_long >113 & c_long <154 & c_lat > (-44) & c_lat < (-10)))
length(unique(df_temp$customer_id))
```
### check the distribution of missing values
```{r}
apply(df, 2, function(x) sum(is.na(x)| x == ''))
# check the number of unique values for each column
apply(df, 2, function(x) length(unique(x)))

```
### filtering out purchase transactions only
```{r}
# assuming purchase transactions must be associated with a merchant (have a merchant Id)
df_temp <- df %>% filter(merchant_id != '' )
# it turned out that is equivilent to excluding following categories of transactions
df_csmp <- df %>%filter(!(txn_description %in% c('PAY/SALARY',"INTER BANK", "PHONE BANK","PAYMEN
T")))
summary(df_csmp)
```
### visualise the distribution of transaction amount
```{r}
hist(df_csmp$amount[!df_csmp$amount %in% boxplot.stats(df_csmp$amount)$out], #exclude outliers
     xlab= 'Transaction Amount', main = 'Histogram of purchase transaction amount')

hist(df$amount[!df$amount %in% boxplot.stats(df$amount)$out], #exclude outliers
     xlab= 'Transaction Amount',main = 'Histogram of overall transaction amount')

df2 <- df %>%
  group_by(customer_id) %>%
  summarise(mon_avg_vol = round(n()/3,0))
hist(df2$mon_avg_vol,
     xlab= 'Monthly transaction volume', ylab='No. of customers', main = "Histogram of customer
s' monthly transaction volume")

```
### Visualise transaction volume over an average week.
```{r}
df3 <- df %>%
  select(date,weekday) %>%
  group_by(date,weekday) %>%
  summarise(daily_avg_vol = n()) %>%
  group_by(weekday) %>%
  summarise(avg_vol=mean(daily_avg_vol,na.rm=TRUE ))
df3$weekday <- factor(df3$weekday, levels=c( "Monday","Tuesday","Wednesday",
                                             "Thursday","Friday","Saturday","Sunday"))
ggplot(df3,aes(x=weekday, y=avg_vol)) +geom_point()+geom_line(aes(group = 1))+
  ggtitle('Average transaction volume by weekday') +
  labs(x='Weekday',y='Transaction volume')

```
### visualize transaction volume over an average week.
```{r}
df4 <- df %>%
  select(date,hour) %>%
  group_by(date,hour) %>%
  summarize(trans_vol=n()) %>%
  group_by(hour) %>%
  summarize(trans_vol_per_hr = mean(trans_vol,na.rm=TRUE))
ggplot(df4,aes(x=hour,y=trans_vol_per_hr))+geom_point()+geom_line(aes(group = 1))+
  ggtitle('Average transaction volume by hour') +
  labs(x='Hour',y='Transaction volume') + expand_limits( y = 0)
```

```{r}
# exclude the single foreign customer whose location information was incorrectly stored (i.e latitude 573)
df_temp <- df_csmp %>%
  filter (c_long >113 & c_long <154 & c_lat > (-44) & c_lat < (-10))
dfloc = df_temp [,c("c_long", "c_lat","m_long", "m_lat")]
dfloc<- data.frame(sapply(dfloc, as.numeric))
dfloc$dst <- distHaversine(dfloc[, 1:2], dfloc[, 3:4]) / 1000
hist(dfloc$dst[dfloc$dst<100], main = "Distance between customer and merchants",xlab= 'Distance
(km)' )

```
